What C++ architecture could best support QoE-oriented map streaming in a player-building-based multiplayer game?

### 2018-03-20 draft.tex
Updated and working through self-review. Let's recover some of those lost marks!!

@@

\section{Introduction}
\begin{figure}
\includegraphics[width=0.75\linewidth]{Newzoo_2017_Global_Games_Market.png}
\caption{The games market in 2017, according to NewZoo}
\label{fig:market}
\end{figure}
The games market is continuing to grow at an immense rate, especially in developing countries. In East Asia, a growth of nearly 15\% over the last year has been observed.\cite{fig:market}

This presents exciting creative opportunities for game developers to expand into new territories. However, a lag in technological advancement, particularly broadband, raises reasonable concerns of whether they can fully exploit these opportunities. The potential of games where players can build the map, such as the extremely popular (as suggested by Google Trends\cite{minecraftnite}) Minecraft and Fortnite, may be hindered by a lack of bandwidth necessary to download player-built resources at an acceptable speed. This may frustrate players by rendering them unable to access areas of the game in a timely manner.

This study aims to discover the extent of this problem, followed by proposals of architectural methods to implement systems that could mitigate the negative effects on the player experience in an unreliable networking environment: This largely comprises of QoE (quality-of-experience) improvements. The programming language in question is C++, while the hypothetical game assumed a 3D environment wherein the player can manipulate the terrain and build meshes. The goal is to propose a reasonable set of design patterns to be used for QoE-oriented map streaming in a large-scale game, and how they could be applied to an existing game engine.

@@

\section{The wider market}
China is in fact the most profitable games market as of 2017 \cite{chinamarket}. Furthermore, in March 2018, 8 of the 10 most popular Steam games in China were partly or fully multiplayer \cite{steamchina}. This suggests China is a valuable target for online games, yet its Internet speeds lag behind (figuratively speaking) the world, at an average of 7.6Mbps \cite{webspeeds}.

Another country notorious for its online gaming community is Brazil, whose Internet speeds averaged about the same--6.8Mbps--in 2017.\cite{webspeeds}. While a marked improvement is evident (doubling since 2015 \cite{webspeeds2015}), to the extent that low-bandwidth games such as FPSes are comfortably playable [cite?], this low speed closes doors to potential large-scale player-built online games.

(Internet speeds in Brazil are notoriously low--an unfortunate feat for the fifth top source of DDoS attacks in early 2017\cite{websecurity}--)

\section{Player experience}
Video streaming is a heavily researched area; perhaps more than games, owing to greater dedicated R\&D toward streaming services such as the Amazon FireTV[cite?]. Considering the aesthetic link between video quality and quality of gameplay--including pixel resolution, lag spike rate, and frame rate--we can draw some preliminary conclusions about a player's interests from studies.

@@

A smooth experience is necessary to retain game players. Studies have noted latency \cite{qossensitivity} \cite{lagragequits}, jitter and packet loss \cite{lagragequits} to all have negative impacts, sometimes leading to quitting altogether. While this is not widely studied in building games--but FPSes and MMOs--games such as the hugely popular (cite?) Fortnite, which features building elements with shooting gameplay, suggest that there may be a demand for player building in those contexts regardless. The potential loss of players therefore justifies the development of strategies to improve the player experience on an unreliable network, especially considering the high traffic potential of level streaming.

>>>>

\newcommand{\source}[1]{\caption*{Source: {#1}} }
% Above code sourced from Xavi, Stack Overflow, https://tex.stackexchange.com/questions/95029/add-source-to-figure-caption @ 20/03/2018

@@

\section{Introduction}
The games market is continuing to grow at an immense rate, especially in developing countries. In East Asia, a growth of nearly 15\% over the last year has been observed, as shown in Fig. \ref{fig:market}.

\begin{figure}[H]
	\includegraphics[width=0.75\linewidth]{Newzoo_2017_Global_Games_Market.png}
	\caption{The games market in 2017, according to NewZoo (Source: \cite{globalmarketpic})}
	\label{fig:market}
\end{figure}

This presents profitable opportunities for game developers to expand into new territories. However, a lag in technological advancement, particularly broadband, raises reasonable concerns of whether they can fully exploit these opportunities. The potential of games where players can build the map, such as the extremely popular (as suggested by Google Trends \cite{minecraftnite}) Minecraft and Fortnite, may be hindered by a lack of bandwidth necessary to download player-built resources at an acceptable speed. A slow download of map resources may frustrate players by rendering them unable to access areas of the game in a timely manner.

This study aims to discover the extent of this problem, followed by proposals of architectural methods to implement systems that could mitigate the negative effects of an unrealiable networking environment on the player experience. These can be described as QoE (quality-of-experience) improvements where the bandwidth is not improved, but the specific targeting of resources is. The programming language in question is C++, while the hypothetical game assumed a 3D environment wherein the player can manipulate the terrain and build meshes. The goal is to propose a reasonable set of design patterns to be used for QoE-oriented map streaming in a large-scale game, and how they could be applied to an existing game engine.

@@

\section{The Wider, Slower Market}
China is the most profitable games market as of 2017 \cite{chinamarket}. Furthermore, according to SteamSpy in March 2018, 8 of the 10 most popular Steam games in China were partly or fully multiplayer \cite{steamchina}. This suggests China is a valuable target for online games. However, compared to most developed countries, China's Internet speeds lag behind (figuratively speaking), at an average of 7.6Mbps \cite{webspeeds}. This inspires concerns that the potential of the multi-player gaming industry, particularly in asset-heavy games like building games, could be hindered in this potentially profitable region.

Similarly, another country notorious for its online gaming community is Brazil. According to SteamSpy, 9 of the top 10 games had a multiplayer component, with 8 boasting primarily multiplayer gameplay \cite{steambrazil}. Brazil's Internet speed average about the same--6.8Mbps--in 2017 \cite{webspeeds}. While a marked improvement over time is evident (doubling since 2015 \cite{webspeeds2015}), to the extent that low-bandwidth games such as FPSes are comfortably playable [cite?], it puts to question whether the development of large-scale player-built online games is halted by this lack of gross speed (an ironic failing for the world's fifth top source of DDoS attacks in early 2017 \cite{websecurity}).

There is a lack of definitive research of the impact of low broadband speed in games. Instead, many studies such as [CITE] are focused on reducing and optimising traffic. This is helpful to developers, but lacks a point of reference or scale, making it plausible that a developer could fear the unknown bottlenecks of slow broadband. Until this is further understood, it could be thus beneficial to design a system around scalable broadband speed, such that it can guarantee an acceptably 'fun' quality even in extremely slow network conditions. However, the definition of 'fun' may be at odds with the industry's tendency to create games that are bigger, more detailed, and overall more data-heavy as they evolve. (CITE) This prompts an exploration into what is 'fun', in a term more aptly described as quality of experience.

\section{Quality of player experience}
Video streaming is a heavily researched area; perhaps more than games, owing to greater dedicated R\&D toward streaming services such as the Amazon FireTV[cite?]. Considering the link between video quality and quality of gameplay aesthetic--including pixel resolution, lag spike rate, and frame rate--some preliminary conclusions about a player's interests can be drawn from this research.

@@

This smooth experience is also necessary to retain game players. Studies have noted latency \cite{qossensitivity} \cite{lagragequits}, jitter and packet loss \cite{lagragequits} to all have negative impacts, sometimes leading to quitting altogether. While this is not widely studied in building games--but FPSes and MMOs--games such as the recent and hugely popular (cite?) Fortnite, which feature both building elements and shooting gameplay, suggest that there may be new demand for player building in those contexts. The potential loss of players therefore justifies the development of strategies to improve the player experience on an unreliable network, especially considering the high traffic potential of level streaming (CITE THIS!).

\section{Streaming costs}
A hypothetical budget map of descending streaming costs follows:
- Textures
- Leel geometry
- Player animations
- LOOK AT A GAME'S MEMORY BUDGET
